---
title: "Assignment 2"
author: "Mayuree"
date: '2022-06-15'
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---
#### Section 1: Data Description

The Hatecrime Dataset  entails the details of the people who died due to Hate crime. It represents Median_household_Income which represents Median household Income,2021.It also represents Share of the population that is unemployed (seasonally adjusted) & metropolitan areas.Also, Share of adults 25 and older with a high-school degree & Share of the population that are not U.S. citizens
Research Question is- Whether High Rates of Crimes is dependent to Income Inequality. Is there a correlation between median_household_income & hate_crimes_per_100k_splc
Format of the File.This data is saved in CSV file format in github. This is a flat file. This is a delimited file & comma is the delimiter used in the file.



#### Section 2: Reading the data into R
```{r, echo = TRUE}
url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/hate-crimes/hate_crimes.csv"
data <- read.csv(url)
```

#### Section 3: Clean the data

#### Renaming column names with upper case.

```{r, echo = TRUE}
library(tidyverse)
data <- data %>%
  rename_with(toupper)
```

#### Renaming columns to Hate_Crime and AVg_HateCrimes 
```{r, echo=TRUE}
data %>% 
  rename(
    Hate_Crime = HATE_CRIMES_PER_100K_SPLC,
    Avg_HateCrimes = AVG_HATECRIMES_PER_100K_FBI
  )

```

#### Section 4: Characteristics of the data
This data frame has rows `r nrow(data)` and columns `r ncol(data)`. The names of the columns and a brief description of each are in the table below:

```{r, echo = TRUE}
Fields <- data.frame(
                               `Column Name` = c('state',	'median_household_income',	'share_unemployed_seasonal','share_population_in_metro_areas',	
                                                 'share_population_with_high_school_degree',	'share_non_citizen',	'share_white_poverty','gini_index',	
                                                 'share_non_white','share_voters_voted_trump','hate_crimes_per_100k_splc','avg_hatecrimes_per_100k_fbi'),
                               
                                Description =  c('State name',
                                                 'Median household income, 2016',
                                                 'Share of the population that is unemployed (seasonally adjusted), Sept. 2016',
                                                 'Share of the population that lives in metropolitan areas, 2015',
                                                 'Share of adults 25 and older with a high-school degree, 2009',
                                                 'Share of the population that are not U.S. citizens, 2015',
                                                 'Share of white residents who are living in poverty, 2015',
                                                 'Gini Index, 2015','Share of the population that is not white, 2015',
                                                 'Share of 2016 U.S. presidential voters who voted for Donald Trump',
                                                 'Hate crimes per 100,000 population, Southern Poverty Law Center, Nov. 9-18, 2016',
                                                 'Average annual hate crimes per 100,000 population, FBI, 2010-2015'
)
)
          
library(knitr)                     
kable(Fields)                           
```

#### Section 5: Summary Statistics
##### missing values
```{r echo=TRUE}
# picking 3 columns from the data set
data_filtered <- data[, c('STATE', 'MEDIAN_HOUSEHOLD_INCOME', 'SHARE_UNEMPLOYED_SEASONAL')]

missing <- as.data.frame(lapply(data_filtered, function(x) sum(is.na(x))))

missing <- rename(missing, state_missing_val = STATE, median_household_income_missing_val = MEDIAN_HOUSEHOLD_INCOME, share_unemployed_seasonal_missing_val = SHARE_UNEMPLOYED_SEASONAL)

missing
```

```{r, echo = TRUE}
# how to create an object
summary_statistics <- summary(data_filtered)
summary_statistics
```


